---
title: "Group20"
author: "314963810_313177073_316088525"
date: "28 4 2021"
output:
word_document: default
 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
 

```
###### Simulation Project

### Libraries


we will need the following libraries: fitdistrplus, magrittr
```{r libraries, include=FALSE,echo=FALSE}
library(fitdistrplus)
library(magrittr)
```
## 1.  Empirical Data

Empirical data - observations from the excel file added to excercise.
We manipulated the data: the time between the intsances (in minutes) was calculated, and we took 1/that time to get the rate samples.

```{r excel1 samples}

samples <- read.csv(file.choose(), header = T)
print(samples)
large_rate <- samples[1:50,5]
mugbalut_rate <- samples[1:50,6]
print(large_rate)
print(mugbalut_rate)
```
#### Large Kalpi

## 2. First Impressions 

We will look at the data and understand - 

1) The data range - **(0, 30)**

2) Is it discrete or continous? **CONTINOUS, A-SYMMETRIC DISTRIBUTION WITH RIGHT TAIL. It looks like an exponental distribution** 

3) are there any outliers? **YES, 3**


We used the following plots to answer the questions above:

```{r empirical plots1, include=TRUE}
par(mfrow=c(1,3)) # defines 2 plots on the same window
plot(large_rate,ylab="Sample values") # plot ( left graph )
hist(large_rate, ylab="Frequency ") # histogram (right graph)
bp1 <- boxplot(large_rate)
```
## 3. Exceptional analysis

We will delete exeptional data with the help of the code below:

```{r exeptional1,include=TRUE}
min_exception1 <- min(bp1$out)
final_large_rate <- subset(large_rate, large_rate < min_exception1)
```

Now we can look further, and plot the histogram not in original values but as density (empirical f(x)) and cummulative probability (F(x)), using plotdist from the fitdistrplus library. 

## 4. Fit a known distribution 

We can use R to estimate the paramters of known distribution to our data.

Notice that it **CAN'T** be the following distibutions:

- beta - because the values are not in [0,1]

- pois/nbinom/geom - because the continous distribution.

The estimate are R estimates of the parameters of a tested distribution - **If it was the exact distribution**.

We will try fitting the model to: exp, norm and gamma distributions.

```{r fitting1, include=TRUE}
expFit<-fitdist(final_large_rate,"exp") # fitting a exp distrbution 
summary(expFit) # summary function displays the results

normFit<-fitdist(final_large_rate,"norm")
summary(normFit)

gammaFit<-fitdist(final_large_rate,"gamma")
summary(gammaFit)

```
We can see the difference in the likelihood and in the BIC/AIC. We want higher likelihood, and lower AIC/BIC. 
**We can conclude that the gamma distribution with shape=1.9688020 and  rate=0.6994063 fits our data better.** 

This will be seen from the plots as well:

```{r individual graphs1, include=TRUE}
plotdist(final_large_rate,"gamma",para=list(shape=gammaFit$estimate[1], rate=gammaFit$estimate[2]))
plotdist(final_large_rate,"exp",para=list(rate=expFit$estimate[1]))
plotdist(final_large_rate,"norm",para=list(normFit$estimate[1],normFit$estimate[2])) 

```
We can see that this is absolutely NOT a norm distribution.
There is similarity between exponental and gamma distributions, **BUT gamma distribution fits much better**.

The two graphs can be combined into one graph in the following manner:
```{r combined graphs1, include=TRUE}
par(mfrow=c(2,2)) # manually define a 2X2 graph window 
legendText<-c("gamma","exp") # define legend text
denscomp(list(gammaFit,expFit),legendtext=legendText) # histograms
qqcomp(list(gammaFit,expFit),legendtext=legendText)  # qq plot
cdfcomp(list(gammaFit,expFit),legendtext=legendText) # cdf plot
ppcomp(list(gammaFit,expFit),legendtext=legendText) # pp plot
```

## 5. Perform Goodness of Fit tests on fitted distribution

Now that we found a fitting distribution, we need to statisticlly test the hypothesis our data came from this distribution. for that we can use Goodness of Fit tests (GOF). We can choose between Chi-Squared (CHISQ) or Kolmogorov-Smirnov (KS).

**H0: arrival~gamma(1.9688020, 0.6994063)**

**H1: NOT**

We will use the general function gofstat with KS because it's more accurate:


```{r, goodness1 KS}
gammagof<- gofstat(gammaFit)%>%print()
gammagof$kstest%>%print()
```

The gamma distribution test - **H0 is not denied**, and hence we can conclude the **data came from gamma distribution**.

p-value for the test is: 
```{r p-value1, results='asis'}
paste(gammagof$chisqpvalue)
```

## Mugbalut Kalpi

## 2. First Impressions 

We will look at the data and understand - 

1) The data range - **(0, 6)**

2) Is it discrete or continous? **CONTINOUS, A-SYMMETRIC DISTRIBUTION WITH RIGHT TAIL. It looks like an exponental distribution** 

3) are there any outliers? **YES, 6**


We used the following plots to answer the questions above:

```{r empirical plots2, include=TRUE}
par(mfrow=c(1,3)) # defines 2 plots on the same window
plot(mugbalut_rate,ylab="Sample values") # plot ( left graph )
hist(mugbalut_rate, ylab="Frequency ") # histogram (right graph)
bp2 <- boxplot(mugbalut_rate)
```
## 3. Exceptional analysis

We will delete exeptional data with the help of the code below:

```{r exeptional2,include=TRUE}
min_exception2 <- min(bp2$out)
final_mugbalut_rate <- subset(mugbalut_rate, mugbalut_rate < min_exception2)
```

Now we can look further, and plot the histogram not in original values but as density (empirical f(x)) and cummulative probability (F(x)), using plotdist from the fitdistrplus library. 

## 4. Fit a known distribution 

We can use R to estimate the paramters of known distribution to our data.

Notice that it **CAN'T** be the following distibutions:

- beta - because the values are not in [0,1]

- pois/nbinom/geom - because the continous distribution.

The estimate are R estimates of the parameters of a tested distribution - **If it was the exact distribution**.

We will try fitting the model to: exp, unif and gamma distributions.

```{r fitting2, include=TRUE}
expFit2<-fitdist(final_mugbalut_rate,"exp") # fitting a exp distrbution 
summary(expFit2) # summary function displays the results

normFit2<-fitdist(final_mugbalut_rate,"norm")
summary(normFit2)

gammaFit2<-fitdist(final_mugbalut_rate,"gamma")
summary(gammaFit2)

```
We can see the difference in the likelihood and in the BIC/AIC. We want higher likelihood, and lower AIC/BIC. 
**We can conclude that the gamma distribution with shape=2.023509 and  rate=2.835970 fits our data better.** 

This will be seen from the plots as well:

```{r individual graphs2, include=TRUE}
plotdist(final_mugbalut_rate,"gamma",para=list(shape=gammaFit2$estimate[1], rate=gammaFit2$estimate[2]))
plotdist(final_mugbalut_rate,"exp",para=list(rate=expFit2$estimate[1]))
plotdist(final_mugbalut_rate,"norm",para=list(normFit2$estimate[1],normFit2$estimate[2])) 

```
We can see that this is absolutely NOT a unif distribution.
There is similarity between exponental and gamma distributions, **BUT gamma distribution fits much better**.

The two graphs can be combined into one graph in the following manner:
```{r combined graphs2, include=TRUE}
par(mfrow=c(2,2)) # manually define a 2X2 graph window 
legendText2<-c("gamma","exp") # define legend text
denscomp(list(gammaFit2,expFit2),legendtext=legendText2) # histograms
qqcomp(list(gammaFit2,expFit2),legendtext=legendText2)  # qq plot
cdfcomp(list(gammaFit2,expFit2),legendtext=legendText2) # cdf plot
ppcomp(list(gammaFit2,expFit2),legendtext=legendText2) # pp plot
```

## 5. Perform Goodness of Fit tests on fitted distribution

Now that we found a fitting distribution, we need to statisticlly test the hypothesis our data came from this distribution. for that we can use Goodness of Fit tests (GOF). We can choose between Chi-Squared (CHISQ) or Kolmogorov-Smirnov (KS).

**H0: arrival~gamma(2.023509, 2.835970)**

**H1: NOT**

We will use the general function gofstat with KS because it's more accurate:


```{r, goodness2 KS}
gammagof2<- gofstat(gammaFit2)%>%print()
gammagof2$kstest%>%print()
```

The gamma distribution test - **H0 is not denied**, and hence we can conclude the **data came from gamma distribution**.

p-value for the test is: 
```{r p-value2, results='asis'}
paste(gammagof2$chisqpvalue)
```


